Hippocampal segmentation with MAGeT

### 2012-12-05 paper.Rnw
Fleshed out methods a bit more.. worked in the rationale

\documentclass{article}
\usepackage[margin=1.5in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
\usepackage[square,numbers]{natbib}
\setlength{\marginparwidth}{1.2in}
\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
\newcommand{\todo}[1]{\marginnote{\textcolor{red}{TODO #1}}}
\newcommand{\mc}[1]{\todo{(MC) #1}}

@@

\section{Materials and Methods}

explain two registration methods, definitions, fusion methods. 
compare to each other, and then compare atlas-based, true multi-atlas, naive methods.
trying to make choices as to the best reg/fusion/

evaluations
described previously in Chakravarty2012.

\subsection{Segmentation Algorithm}
\todo{Make sure we explain what MAGeT stands for}
In this paper, we use the term {\it atlas} to mean any manually segmented MR
image, and the term {\it atlas library} to mean a set of such images.  We use
the term {\it template} to refer to any MR image, and associated labelling,
used to segment another image, and the term {\it template library} to refer to
a set of such images.  An atlas library may be used as a template library but,
as we will discuss, a template library may also be composed of images with
computer generated labellings.

@@

Label fusion is performed by cross-correlation weighted voting, a strategy
weighted towards an optimal combination of subjects from the template library
which has been previously shown to improve segmentation
accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
image is ranked in similarity to each unlabelled image by the normalized
cross-correlation of image intensities after linear registration in a region of
interest (ROI) generously encompassing the hippocampus.  Only the top ranked
template library image labels are used in a voxel-wise majority vote. The ROI
is heuristically defined as the extent of all atlas labels after linear
registration to the template, dilated by three voxels \citep{Chakravarty2012}.

@@

\begin{algorithm}
\caption{Pseudocode for the MAGeT Brain algorithm}
\label{pseudocodesdf}
\begin{algorithmic}
\Function{MultiAtlas}{Templates, Subjects}
  \ForAll{$subject$}
    \ForAll{$template$}
      \State propagate all labels for template to subject space
      \State store subject labels
    \EndFor
    \State fuse subject labels
  \EndFor
\EndFunction
\\
\Function {MAGeTBrain}{Subjects, Atlases, n}
  \For{$i = 1 \to n$}
    \State choose a subject to be used as a template
    \State propagate labels from each atlas to template space
    \State store the template with all of its labels
  \EndFor
  \State MultiAtlas(Templates, Subjects)
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\subsection{Image Processing and Registration Methods}

Before images were registered, the N3 algorithm [16] is first used to minimize
the intensity nonuniformity in each of the atlases and unlabeled subject
images. In this work we investigated the performance of two registration
methods.

\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}

The ANIMAL algorithm carries out Image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the
whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
deformation field between two MR images. At first, large deformations are
estimated using blurred version of the input data. These larger deformations
are then input to subsequent steps where the fit is refined by estimating
smaller deformations on data blurred with a Gaussian kernel with a smaller
FWHM. The final transformation is a set of local translations defined on a bed
of equally spaced nodes that were estimated through the optimization of the
correlation coefficient.

For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
should be noted that the MAGeT brain algorithm is not dependent on this, or
any, particular choice of registration method \citep{Chakravarty2012}.

\todo{should add the MNI website here; also we should add the parameters we used somewhere}

@@

ANTs is a diffeomorphic registration algorithm which provides great flexibility
over the choice of transformation model, objective function, and the
consistency of the final transformation. The transformation is estimated in a
hierarchical fashion where the MRI data is subsampled, allowing large
deformations to be estimated and successively refined at later hierarchical
stages (where the data is subsampled to a finer grid). The deformation field
and the objective function are regularized with a Gaussian kernel at each level
of the hierarchy. The ANTs algorithm is freely available
\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the ANTS
algorithm compatible with the MINC data format, mincANTS
\url{https://github.com/vfonov/mincANTS} \todo{(same here we should add the
parameters we used).}

\subsection{Label Fusion Methods}

@@

Labels are propagated from only those template library images which most
similar to the subject.  A voxel-wise majority vote is carried out from the
labels from the top n ranked template library images, where n is parameter set
by the user.

Similarity between a subject and a template image is measured using normalized
cross-correlation of intensity values, over a region of interest defined on
each template, after the subject has been linearly registered to the template.
The heuristic used for defining a meaningful region of interest for each
template is to use the extent of all the propagated labellings from each atlas
(after linear registration only) and then dilating this region by three voxels.

The utility to calculate the cross-correlation similarity measure is
implemented as part of the ANIMAL toolkit. \todo{(ANIMAL or MINC)}

@@

This label fusion method is identical to the above process except a normalised
mutual information score is used over the region of interest between a template
and subject instead of cross-correlation.

The utility to calculate the normalised mutual information measure is
implemented as part of the EZMinc package (based on ITK NMI routine).

@@

\subsection{Data and Experiments}

Three different experiments were conducted to evaluate the influence of
registration method, atlas set size and resolution, template set size, and
label fusion method and parameters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                         Experiment Rationale 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\subsubsection{Study Method Rationale - An Outline}
\begin{description}
\item[ADNI Repeated Random Subsampling Validation] \hfill \\ 
{\bf Purpose:} to show method leads to improved accuracy (over multi-atlas method), and is robust
w.r.t. registration method, atlas and template library composition and size, and various fusion methods. Additionally, to choose the best configuration of parameters (under the assumption that performance generalises). 
\\ \\ 
{\bf Method:} using a set of 69 ADNI subjects, 13 from each disease category, perform 
10 rounds of repeated random subsampling to pick atlas library and template library 
for each parameter setting and then run MAGeT brain to segment non-atlas subjects.
Also perform multi-atlas and single-atlas segmentation using the same parameter settings
and atlas library size.
\\ \\
{\bf Results:} find significant improvement over multi-atlas performed with the same
parameters. Also, find smoothed performance is monotonically increasing but asymptotic
in size of both template and atlas library, with peak performance reached after 15
templates. \todo{Can we statistically capture "peak" performance?  The point at which 
gains become statistically insignificant?}

\item[Winterburn-Atlas ADNI Baseline Segmentation] \hfill \\
{\bf Purpose:} given that MAGeT is robust and accurate, show MAGeT performance on 
a real-world task in which atlas set may not be representative of subjects.
\\ \\
{\bf Method:} We use best settings from first experiment (ANTS, majority vote, 
template library of 20), and use Winterburn hi-res atlases to segment all ADNI Baseline 
images.  Since our segmentation protocol differs from what can be segmented at 
ADNI resolution a direct comparison is not appropriate.  Instead, we validate by 
showing how well our segmentation volumes correlate with existing automated/manual 
segmentation volumes. \todo{Also show disease prediction?}
\\ \\
{\bf Results: }

\item[First Episode SZ Patients Segmentation] \hfill \\ 
{\bf Purpose:} validate that MAGeT's performance generalises to other diseases.
\\ \\
{\bf Method:} Measure performance using best parameter settings on a different disease dataset. We use FEP SZ patients and controls, using two different atlas sets: a manual hippocampal segmentation of patients, and Winterburn atlas set.  We validate the native-atlas set segmentations using Dice's Kappa, and the Winterburn-atlas-set segmentations by correlating volumes. 
\\ \\
{\bf Results:} High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.

\item[Winterburn Atlas Reliability] \hfill \\
{\bf Purpose:} validation of performance on hippocampal subregion segmentation
\\ \\
{\bf Method:} we use the Winterburn atlas set, using each high-res atlas in turn 
to segment a standard-res MR of the same subject and compare the automatic 
segmentation to the atlas segmentation.  We use a template library consisting 
of standard-res MRs from the remaining Winterburn atlas subjects plus 11 other 
standard-res healthy control subjects. 
\\ \\
{\bf Results:} TBD

\end{description}
\end{kframe}

\subsubsection{ADNI Validation Experiment}

In this experiment we performed repeated random sub-sampling cross-validation
of the MAGeT algorithm with a pool of 69 images and labels from the ADNI
dataset.  The images are T1-weighted \todo{<insert description of scan/machine
types>}.  The hippocampal labels are provided as part of the ADNI dataset, and
are produced \todo{<insert description of SNT labellings>}.  The parameters we
varied, and the ranges over which we varied are listed in \todo{table}.  We
performed 10 validations per parameter combination.  In each validation trial,
a random sample of subjects were chosen from the pool without replacement as
the atlas set, whereas the template set was chosen randomly with replacement.

\begin{table}
    \begin{tabular}{l|l}
        Parameter           & Values tested                                                      \\ \hline
        Number of Templates & 3 to 20                                                            \\ 
        Number of Atlases   & 3 to 9                                                             \\ 
        Registration Method & ANTS or ANIMAL                                                     \\ 
        Label Fusion Method & majority vote, cross-correlation weighted vote, NMI weighted vote. 
    \end{tabular}
\end{table}

We used the following parameters for ANTS, during registration: 
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
These settings were adapted from the "reasonable starting point" given in the
ANTS manual. \todo{<rationale?>}

For the purposes of this work we used the regularization parameters optimized
in Robbins et al.[15]. \todo{right? (but added a second layer).}

\todo{multi-atlas comparison, "naive" comparison}

\subsubsection{Winterburn High-resolution Hippocampal Atlas - ADNI Validation}

In this experiment we explored using the MAGeT brain algorithm with the
Winterburn high-resolution atlases to segment 21 randomly chosen MR images from
the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn
atlases are digital segmentations of the hippocampus in five  in-vivo 300u
isotropic T1-weighted MR scans, and include subfield segmentations for the
cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
Subjects in the Winterburn atlases range in age from 29-57 years (mean age of
37), and include two males and three females.

Since hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct similarity comparisons
between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate
the performance of MAGeT brain, we compared classification accuracy of subjects
by diagnosis based on hippocampal volume using both the SMT labels and our
produced labels.  \todo{description of validation -- rms {\tt validate} or
t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}

<<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c("295","413","619","685","729","782","938","954",
                 "1018","1155","907","981","222","324","448","546",
                 "553","572","602","610","814","1341","675","681",
                 "731","1130","41","68","70","101","249","293","316",
                 "414","698","1206","1222","1339","751","1030","3",
                 "5","10","16","22","53","168","183","241","1205",
                 "328","1095","991","213","159","337","343","642",
                 "753","1109","1217","29","56","1188","1293","149",
                 "382","431","1202")
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTMARRY + PTETHCAT + PTRACCAT + CDRSB + ADAS13 + MMSE, 
               data = adnimerge, subset = VISCODE == 'bl' & RID %in% adni_69_RIDs, 
               method = "reverse", test=TRUE, overall=TRUE)

latex(tab, file="", size="scriptsize", where="h",landscape=TRUE, ctable=TRUE)
@

\todo{For minctracc -- we registered to TAL}

\subsection{Results}
\subsubsection{A2A}

Kappa vs. Number of templates: 
Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
<<a2a_facet, cache=TRUE, fig=TRUE>>=
t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')

@@

ggplot(all_data, aes(x=num_templates, y=k, colour=as.factor(num_atlases))) + 
  facet_grid(method ~ reg_method) + 
  geom_smooth() + 
  geom_hline(data = df_naive, aes(yintercept = value, linetype="dashed")) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + ylab( "Mean Kappa" ) 
@

<<a2a_facet_mameans, cache=TRUE, fig=TRUE>>=
ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 
  facet_grid(method ~ reg_method) + 
  geom_smooth() + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + ylab( "Difference in mean Kappa between voting method and multiatlas" )
@

Kappa minus multi-atlas mean vs. Number of Templates

@@

<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="")
@
\begin{table}
    \begin{tabular}{c|c|c}
        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
    \end{tabular}
\end{table}

@@

\subsubsection{First Episode Schizophrenic patients}

<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ, data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize", where="h")
@

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
\usepackage[square,numbers]{natbib}
\setlength{\marginparwidth}{1.2in}
\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
\newcommand{\todo}[1]{\begin{kframe}{\textcolor{red}{TODO #1}}\end{kframe}}
\newcommand{\mc}[1]{\todo{(MC) #1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                               METHODS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

%%%%%                     MAGeT Brain Algorithm                            %%%%%

\subsection{MAGeT Brain Algorithm}
\todo{Make sure we explain what MAGeT stands for}
\todo{reference Chakravarty2012}
In this paper, we use the term {\it atlas} to mean any manually segmented MR
image, and the term {\it atlas library} to mean a set of such images.  We use
the term {\it template} to refer to any MR image, and associated labelling,
used to segment another image, and the term {\it template library} to refer to
a set of such images.  An atlas library may be used as a template library but,
as we will discuss, a template library may also be composed of images with
computer generated labellings.

@@

@@

\begin{algorithm}
  \scriptsize
  \caption{Pseudocode for the MAGeT Brain algorithm}
  \label{pseudocodesdf}
  \begin{algorithmic}
    \Function{MultiAtlas}{Templates, Subjects}
      \ForAll{$subject$}
        \ForAll{$template$}
          \State propagate all labels for template to subject space
          \State store subject labels
        \EndFor
        \State fuse subject labels
      \EndFor
    \EndFunction
    \\
    \Function {MAGeTBrain}{Subjects, Atlases, n}
      \For{$i = 1 \to n$}
        \State choose a subject to be used as a template
        \State propagate labels from each atlas to template space
        \State store the template with all of its labels
      \EndFor
      \State MultiAtlas(Templates, Subjects)
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\todo{Note that algorithm does not specify a registration algorithm or fusion 
method\citep{Chakravarty2012}.  Performance on these is test in experiments}

%%%%%                             Subjects                                 %%%%%

%  Describe subject datasets in general 
\subsection{Subjects}

\subsubsection{ADNI-1 1.5T Screening}
\todo{- image characteristics}
\todo{- describe SNT manual segmentations}
\todo{- baseline/cross-validation dataset demographics?}

<<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
               data = adnimerge, subset = VISCODE == 'bl',   
               method = "reverse", test=FALSE, overall=TRUE)

latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
      caption="ADNI-1 1.5T Screening demographics")
@
\todo{Check table only includes 1.5T from the Screening}

\subsubsection{SZ First Episode Patients}
\todo{image characteristics, demographics}
<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
               data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize", 
      caption="Schizophrenia First Episode Patient Demographics")
@

%%%%%                       Image pre-processing                           %%%%%

\subsection{Image pre-processing}

Before images were registered, the N3 algorithm [16] is first used to minimize
the intensity nonuniformity in each of the atlases and unlabeled subject
images.

%%%%%                       Registration Methods                           %%%%%

\subsection{Registration}

\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}
The ANIMAL algorithm carries out Image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the
whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
deformation field between two MR images. At first, large deformations are
estimated using blurred version of the input data. These larger deformations
are then input to subsequent steps where the fit is refined by estimating
smaller deformations on data blurred with a Gaussian kernel with a smaller
FWHM. The final transformation is a set of local translations defined on a bed
of equally spaced nodes that were estimated through the optimization of the
correlation coefficient.

For the purposes of this work we used the regularization parameters optimized 
in Robbins et al. \citep{Robbins2004}.

\todo{link to MNI website here}
\todo{command line}

@@

ANTs is a diffeomorphic registration algorithm which provides great flexibility
over the choice of transformation model, objective function, and the
consistency of the final transformation. The transformation is estimated in a
hierarchical fashion where the MRI data is subsampled, allowing large
deformations to be estimated and successively refined at later hierarchical
stages (where the data is subsampled to a finer grid). The deformation field
and the objective function are regularized with a Gaussian kernel at each level
of the hierarchy. The ANTs algorithm is freely available
\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the ANTS
algorithm compatible with the MINC data format, mincANTS
\url{https://github.com/vfonov/mincANTS}.

We used the following command line when running the ANTS command, 
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
These settings were adapted from the "reasonable starting point" given in the
ANTS manual. \todo{reference manual?}

%%%%%                           Label Fusion                               %%%%%

\subsection{Label Fusion}

@@

An optimal combination of subjects from the template library has previously been
shown to improve segmentation accuracy \citep{Aljabar2009,Collins2010}.  In this
method, each template library image is ranked in similarity to each unlabelled 
image by the normalized cross-correlation (CC) of image intensities after linear
registration, over a region of interest (ROI) generously encompassing the 
hippocampus.  Only the top ranked
template library image labels are used in a voxel-wise majority vote. The ROI
is heuristically defined as the extent of all atlas labels after linear
registration to the template, dilated by three voxels \citep{Chakravarty2012}.  
The number of top ranked template library image labels is a configurable parameter.

The {\tt xcorr\_vol utility} from the ANIMAL toolkit is used to calculate the 
cross-correlation similarity measure.

@@

This method is similar to cross-correlation weighted voting 
except that image similarity is calculated by the normalised 
mutual information score over the region of interest. \todo{reference?}

The {\tt itk\_similarity} utility from the EZMinc toolkit is used to calculate 
the normalised mutual information measure.

%%%%%                        Goodness-of-fit                               %%%%%

@@

%%%%%                            Experiments                               %%%%%

\subsection{Experiments}

Experiments were performed to assess the performance of MAGeT brain with various 
parameter settings as well as on diverse datasets.  In each experiment we contrast 
the performance of MAGeT brain with that standard single- and multi-atlas 
segmentations derived from the same atlas library.

%%%%%
%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
%%%%%
\subsubsection{ADNI-1 cross-validation}

To test the accuracy of the MAGeT brain algorithm with different parameter 
settings, repeated random sub-sampling cross-validation (RRSCV) was performed
on a subset of the ADNI-1 dataset.

{\bf Dataset evaluated.} 69 1.5T images were randomly selected from the 
{\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
are shown in Table X\todo{FIXME: table reference}.

<<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c("295","413","619","685","729","782","938","954",
                 "1018","1155","907","981","222","324","448","546",
                 "553","572","602","610","814","1341","675","681",
                 "731","1130","41","68","70","101","249","293","316",
                 "414","698","1206","1222","1339","751","1030","3",
                 "5","10","16","22","53","168","183","241","1205",
                 "328","1095","991","213","159","337","343","642",
                 "753","1109","1217","29","56","1188","1293","149",
                 "382","431","1202")
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
               data = adnimerge, subset = VISCODE == 'bl' & RID %in% adni_69_RIDs, 
               method = "reverse", test=FALSE, overall=TRUE)

latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
      caption="ADNI-1 cross-validation subset demographics")
@

{\bf Atlas and template libary.}  Atlases consisted of images taken from the 
dataset, with corresponding manual labels provided by SNT.  Atlas library size
was varied from 3 to 9 images.  The remaining images were segmented, with the 
template library size varying from 3 to 20 images.  Template library images were
selected randomly from the images to be segmented.

{\bf Registration method.}  Both the ANTS and ANIMAL registration methods were 
used.
\todo{In ANIMAL, images aligned to TAL space}

{\bf Label fusion.} Majority vote, Cross-correlation weighted majority vote, and
Normalized Mutual Information weighted majority vote are used.  With the 
weighted majority vote fusion methods, the number of top labels used in the 
fusion was varied from 3 to 20 images.

{\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV) 
consists of repeated trials in which items from the dataset are randomly assigned to a training set or validation set. In each trial, performance on the validation set is measured, and then averaged across all trials.

We performed RRSCV on each combination of parameters listed above: atlas library size, template library size, registration method, and label fusion method.  We performed 10 trials per parameter combination.  In each validation trial, the training set consisted of the images used as atlases, and the validation set consisted of the images to be segmented. The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were applied to segment the images in the validation set.  Additionally, in each trial, the single-atlas segmentation was obtained for each atlas-template.

The gold-standard for the segmentation accuracy of images in the validation set was the SNT manual labels.

\todo{ number of registrations/comparisions }

%%%%%
%%%%%               Experiments: ADNI-1 Screening Validation               %%%%%
%%%%%
\subsubsection{ADNI-1 Screening Validation}

To test the accuracy of MAGeT brain on a real-world task we segment the entire
ADNI-1 dataset using an atlas set that is not representative of the subject set.

{\bf Dataset evaluated.} All images from the {\em ADNI1:Screening 1.5T} 
standardized dataset.

{\bf Atlas and template libary.} The atlas library consisted of the entire 
Winterburn atlas set.  The Winterburn atlases are digital segmentations of the
hippocampus in five in-vivo 300u isotropic T1-weighted MR scans, and include 
subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, 
subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases range in 
age from 29-57 years (mean age of 37), and include two males and three females.

The template library consisted of 21 randomly selected images from the ADNI1 
data dataset (7 healthy, MCI and AD subjects).

{\bf Registration method.}  - ANTS as it performed best in the cross-validation
experiment. \todo{how to reference results whilst still in Methods?}

{\bf Label fusion.} Majority vote, as it is simpliest to run and performed 
equally well in cross-validation experiment.

{\bf Evaluation.}

Since hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct similarity comparisons
between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
we did(n't) resegment the ADNI images with a the low-res protocol.}

To evaluate the performance of MAGeT brain, we correlate our segmentation
volumes with manual segmentation volumes, as well as with hippocampal volumes
of established automated segmentation methods.

Additionally, we compared classification accuracy of subjects
by diagnosis based on hippocampal volume using both the SMT labels and our
produced labels.  \todo{description of validation -- rms {\tt validate} or
t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}

%%%%%
%%%%%          Experiments: SZ First Episode Patient Validation           %%%%%
%%%%%
\subsubsection{SZ First Episode Patient Validation}
 To validate that MAGeT's performance generalises to other diseases.
Measure performance using best parameter settings on a different disease dataset.

{\bf Dataset evaluated.} SZ First Episode Patients

{\bf Atlas and template libary.} two different atlas sets: a manual hippocampal
segmentation of patients, and Winterburn atlas set.

{\bf Registration method.} ANTS.

{\bf Label fusion.} Majority vote.

{\bf Evaluation.} We validate the FEP-atlas segmentations using Dice's Kappa,
and the Winterburn-atlas segmentations by correlating volumes.

%%%%%
%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
%%%%%
\subsubsection{Winterburn Atlases Validation}
{\bf Dataset evaluated.} T1 BRAVO scans of the same subjects included in the 
Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
the Winterburn atlases.

{\bf Atlas and template libary.} Atlas library is Winterburn T1 atlases.  
Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.

{\bf Registration method.} ANTS

{\bf Label fusion.} Majority vote.

{\bf Evaluation.}  Leave one out cross-validation (LOOCV) in which all five
subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
to be segmented is excluded from the Atlas library (so only four atlases are used).

Segmentation accuracy is judged by difference in hippocampal volume.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                               Results
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results}

%%%%%
%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
%%%%%
\subsubsection{ADNI-1 Cross-Validation}

 - find significant improvement over multi-atlas performed with the same
  parameters. Also, find smoothed performance is monotonically increasing but asymptotic
  in size of both template and atlas library, with peak performance reached after 15
  templates. \todo{Can we statistically capture "peak" performance?  The point at which 
  gains become statistically insignificant?}

\begin{figure}
<<ADNI1_cross_validation_means, cache=TRUE, fig=TRUE>>=
t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')

@@

ggplot(all_data, aes(x=num_templates, y=k, colour=as.factor(num_atlases))) + 
  facet_grid(method ~ reg_method) + 
  geom_smooth() + 
  geom_hline(data = df_naive, aes(yintercept = value, linetype="dashed")) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + ylab( "Mean Kappa" ) 
@
  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))}
  \label{ADNI1_cross_validation_means}
\end{figure}

\begin{figure}
<<ADN1-cross-validation-MA-difference, cache=TRUE, fig=TRUE>>=
ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 
  facet_grid(method ~ reg_method) + 
  geom_smooth() + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + ylab( "Difference in mean Kappa between voting method and multiatlas" )
@
  \caption{Difference in mean Kappa between MAGeT brain and multi-atlas}
  \label{ADNI1_cross_validation_MA_difference}
\end{figure}

@@

<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="", dec=2)
@
\begin{table}
    \begin{tabular}{c|c|c}
        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
    \end{tabular}
\end{table}

@@

\subsubsection{First Episode Schizophrenic patients}
High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
